\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2013-11-03 doc/paper.tex
fleshed out more of the structure.

%macros
\newcommand{\sarah}{SARAH\xspace}
\newcommand{\fs}{FlexibleSUSY\xspace}
\newcommand{\ESSM}{E$_6$SSM\xspace}
\newcommand{\code}[1]{\lstinline|#1|}  % inline source code
\newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}
\newcommand{\DRbar}{\textoverline{DR}}  % inline source code

@@

 The modular structure of the generated code allows for any individual
 components to be replaced with an alternative if available.  Indeed
 \fs has been intentionally designed to grow as alternative solvers
 and calculators are added and a subsequent release with an
 alternative lattice solver is already planned.{\color{red} Or Indeed
   flexible SUSY is distributed with two diffrent main solvers, the
   two-scale solver based omn the approach of \cite{} and designed for
   speed and the a lattice solver designed to solve tricky cases where
   the fixed point iteration of the two-scale approach fails to
   converge.}

@@

\subsection{Quick Start}
Create a MSSM spectrum generator
  \begin{lstlisting}[language=bash]
$ ./createmodel --models=MSSM
$ ./configure --with-models=MSSM
$ make
  \end{lstlisting}
Run the spectrum generator using the default parameter point
  \begin{lstlisting}[language=bash]
$ ./models/MSSM/run_MSSM.x
  \end{lstlisting}
Run the spectrum generator using a SLHA input file
  \begin{lstlisting}[language=bash]
$ ./models/MSSM/run_MSSM.x --slha-input-file=templates/MSSM/LesHouches.in.MSSM --slha-output-file=LesHouches.out.MSSM
  \end{lstlisting}

\section{Stucture of the spectrum generator}
\subsection{boundary conditions}
Different types, how we match to low energy data.  reference EWSB or merge that section here.
\subsection{RGEs}
\subsection{tree level spectrum}
\subsection{Electroweak symmetry breaking}
\subsection{pole masses}
\subsection{two scale fixed point iteration}
\subsection{Lattice method}
If ready.
\subsection{multiple effective models at different scales}
\section{Implementing new models}
\section{Adapting the code at the C++ level}
\appendix
\section{Examples}
\subsection{NMSSM}
Detailed steps to create NMSSM model file.
\subsection{\ESSM}
Detailed steps to create \ESSM model file.

\section{Tests}

@@

>>>>

%macros
\newcommand{\sarah}{SARAH\xspace}
\newcommand{\fs}{FlexibleSUSY\xspace}
\newcommand{\ESSM}{E$_6$SSM\xspace}
\newcommand{\code}[1]{\lstinline|#1|}  % inline source code
\newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}
\newcommand{\DRbar}{\textoverline{DR}}  % inline source code
\def\at{\alpha_t}
\def\ab{\alpha_b}
\def\as{\alpha_s}
\def\atau{\alpha_{\tau}}
\def\oat{O(\at)}
\def\oab{O(\ab)}
\def\oatau{O(\atau)}
\def\oatab{O(\at\ab)}
\def\oatas{O(\at\as)}
\def\oabas{O(\ab\as)}
\def\oatababq{O(\at\ab + \ab^2)}
\def\oatqatababq{O(\at^2 + \at\ab + \ab^2)}
\def\oatasatq{O(\at\as + \at^2)}
\def\oatasabas{O(\at\as +\ab\as)}
\def\oatasabasatq{O(\at\as + \at^2 +\ab\as)}
\def\oatq{O(\at^2)}
\def\oabq{O(\ab^2)}
\def\oatauq{O(\atau^2)}
\def\oabatau{O(\ab \atau)}
\def\oas{O(\as)}
\def\oatauqatab{O(\atau^2 +\ab \atau )}

@@

 The modular structure of the generated code allows for any individual
 components to be replaced with an alternative if available.  Indeed
 \fs has been intentionally designed to grow as alternative solvers
 and calculators are added and a subsequent release with an
 alternative lattice solver is already planned.{\color{red} Or Indeed
   flexible SUSY is distributed with two different main solvers, the
   two-scale solver based on the approach of \cite{} and designed for
   speed and the a lattice solver designed to solve tricky cases where
   the fixed point iteration of the two-scale approach fails to
   converge.}

@@

\subsection{Quick Start}
Create a MSSM spectrum generator
  \begin{lstlisting}[language=bash]
$ ./createmodel --models=MSSM
$ ./configure --with-models=MSSM
$ make
  \end{lstlisting}
Run the spectrum generator using the default parameter point
  \begin{lstlisting}[language=bash]
$ ./models/MSSM/run_MSSM.x
  \end{lstlisting}
Run the spectrum generator using a SLHA input file
  \begin{lstlisting}[language=bash]
$ ./models/MSSM/run_MSSM.x --slha-input-file=templates/MSSM/LesHouches.in.MSSM --slha-output-file=LesHouches.out.MSSM
  \end{lstlisting}%% $

\section{Stucture of the spectrum generator}
\subsection{boundary conditions}
Different types, how we match to low energy data.  reference EWSB or merge that section here.
mention two loop qcd parts from softsusy.
\subsection{RGEs}
The two loop renormalisation group equations are generated using command ?
Even in the simplest supersymmetric model the RGEs form a very large set of coupled non-linear differential equations.  To solve these and thereby evolve the parameters between two different scales \fs has... 
\subsection{tree level spectrum}
The diagonalised tree level $\overline{DR}$ masses are obtained from expressions...
The numerical diagonalisation is performed by... with an accuracy set by...   
\subsection{Electroweak symmetry breaking}
At tree level the EWSB conditions are written...
The one loop electroweak symmetry breaking are solved using the gsl routines finding multiple roots, \code{gsl_multiroot_fsolver}.  Iteration. For example in the MSSM \fs generates.. show code.
 If higher accuracy is required additional routines with higher order corrections can be added in a very simple way.  For example in the MSSM by default \fs adds two loop Higgs FORTRAN routines supplied by P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of $\oatas$, $\oabas$, $\oatq$, $\oabatau$, 
$\oabq$, $\oatauq$ and $\oatab$  as follows...  
Something similar is done for the NMSSM, but in this case the NMSSM $\oatas$, $\oabas$ pieces come from \cite{Degrassi:2009yq}.
 Corrections from other sources can be added to the generated code by...
\subsection{two scale fixed point iteration}
\subsection{pole masses}
With the solver routine finished and convergence achieved, the $\overline{DR}$ parameters consistent with one loop EWSB, low energy data and all user supplied boundary conditions are then known at any scale between...

The physical mass spectrum can now be calculated.  \fs uses the full one-loop self energies and tree level routines to calculate the pole masses at the one-level.  If higher accuracy is required additional routines with higher order corrections can be added in a very simple way. For example in the MSSM by default \fs adds two loop Higgs FORTRAN routines supplied by P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of $\oatas$, $\oabas$, $\oatq$, $\oabatau$, 
$\oabq$, $\oatauq$ and $\oatab$  as follows...  
Something similar is done for the NMSSM, but in this case the NMSSM $\oatas$, $\oabas$ pieces come from \cite{Degrassi:2009yq}, while for $\oatq$, $\oabatau$, 
$\oabq$, $\oatauq$ and $\oatab$ the MSSM pieces are used though it should be understood that these are not complete in the NMSSM. For other models since the Higgs mass is a very important measurement and the two loop corrections can be larger than the current experimental error \cite{Degrassi:2009yq} we recommend that the leading log two loop corrections are estimated, by generalising those of the MSSM or NMSSM, as has been done, for example, in the E$_6$SSM\cite{King:2005jy}.

Corrections to other masses or further corrections to the Higgs states can be added very easily by...

\subsection{Lattice method}
If ready.
\subsection{multiple effective models at different scales}
\section{Implementing new models}
\section{Adapting the code at the C++ level}
\appendix
\section{Examples}
\subsection{NMSSM}
Detailed steps to create NMSSM model file.
\subsection{\ESSM}
Detailed steps to create \ESSM model file.

\section{Tests}
Comparisons against existing spectrum generators to test for deviations, e.g. the tests against MSSM and NMSSM (and E6SSM?).

Comparisons against existing spectrum generators and SARAH generated spheno code for speed.

@@

%\cite{King:2005jy}
\bibitem{King:2005jy} 
  S.~F.~King, S.~Moretti and R.~Nevzorov,
  %``Theory and phenomenology of an exceptional supersymmetric standard model,''
  Phys.\ Rev.\ D {\bf 73}, 035009 (2006)
  [hep-ph/0510419].
  %%CITATION = HEP-PH/0510419;%%
  %130 citations counted in INSPIRE as of 03 Nov 2013
